Experimental and numerical studies of the effects of resonant and random excitations of the proton beam in the LHC, with applications to the design of a pulsed hollow electron lens for active halo control

### 2018-02-27 resex.tex
Intermediate step in revision of experimental section.

\subsection{Overview of LHC experiments\label{sec:exp_sum}}
The experiments at the LHC which are presented in this paper were aiming at deriving tolerances on the residual HEL field at the location of the beam core in case of pulsed operation. In total, two experiments were conducted, the first one in 2016 \cite{resexmd2016} and the second one in 2017 \cite{resexmd2017}. Beam and machine parameters for both experiments are summarized in Table~\ref{tab:md_param}. Preceding the beam experiments, simulations were performed to identify the most effective $k^{\mathrm{th}}$~turn pulsing pattern and yield estimates for losses and emittance growth. The two $k^{\mathrm{th}}$~turn pulsing patterns with the largest effect, one $k^{\mathrm{th}}$~turn pulsing pattern with no effect and the random excitation could then be experimentally tested in the LHC. In order to test the reproducibility of the results at least on the example of one pulsing pattern, the $7^{\mathrm{th}}$~turn pulsing was first tried in 2016 and then repeated in 2017.

During the experiments the losses were measured with the fast beam current transformers (FBCTs) and the emittance and transverse beam profiles with the beam synchrotron radiation telescope (BSRT), where all instruments are capable of delivering bunch-by-bunch data. The analysis of the BSRT profiles is quite involved and we will therefore focus in this paper only on the direct comparison of the profiles. A detailed description of the analysis with more details can be found in \cite{bsrtprofinj} and for the individual experiments in \cite{resexmd2016,resexmd2017}.

\subsection{Excitation with transverse damper and filling scheme\label{sec:adt}}
The primary function of the transverse feedback system in LHC, also referenced to as ADT, is to provide injection oscillation damping and actively damp the coupled bunch instabilities driven by the machine impedance. The main building blocks of the ADT system are stipline pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to the beam position measurement modules at the surface, the digital signal processing modules (mDSPU) and a set of tetrode power amplifiers feeding electrostatic kickers in the RF zone of the LHC~\cite{adt_sum_2008,adt_sum_2011}. Though originally designed to damp the undesired transverse activities only, the damper is being routinely used for sophisticated beam excitations as well thanks to the state of the art digital signal processing hardware. These include among others abort and injection gap cleaning, single bunch excitation for tune and linear coupling measurements and special modes of operation for dedicated experiments at the LHC.

@@

The ADT kicker deflection angle is calculated from the kicker geometry and the excitation voltage on the deflection plates. The excitation voltage used in the experiments is a complicated waveform, generated by a very complex hardware (digital signal processor, chain of transmission lines, low and high power amplifiers). The peak excitation voltage (and subsequently the kick) seen by the beam can be only indirectly measured using the probes on each kicker. The measured values are therefore believed to be known with an error margin of 10-15\% for the experiments in 2017 and 50\% in 2016. The estimate for 2017 is based on only a single measurement and in the future and with more time available the determination of the error would definitely have to be improved. The error in 2016 is estimated from the transverse damper operational parameters. Due to lack of the dedicated machine time no precision calibration was performed. The maximum kick strength reachable with the ADT at injection without risking saturation is 96~nrad. As an example Fig.~\ref{fig:fill_meas} shows the oscillation amplitude of each bunch in both H and V planes captured by the LHC real time transverse activity monitor during the 2017 experiments and an excitation with $8^{\mathrm{th}}$ turn pulsing. The measured activity is well in line with the expected excitation pattern.
\begin{figure*}[h]
	\centering
	\includegraphics[width=1.0\linewidth]{bunchfilling_measured.png}	
	\caption{\label{fig:fill_meas} Bunch by bunch excitation detected by the real time transverse activity monitor during the 2017 experiment and an excitation with 8th turn pulsing.}
\end{figure*}

>>>>

\begin{figure*}
  \includegraphics[width=\textwidth]{bunchfilling_measured.png}	
  \caption{Example of bunch-by-bunch excitation detected by the
    real-time transverse activity monitor during the 2017 experiment,
    when the beam was excited every $8^\mathrm{th}$~turn.}
  \label{fig:fill_meas} 
\end{figure*}

\subsection{Overview}
\label{sec:exp_sum}

The purpose of the experiments at the LHC is to quantify the effects
of a pulsed excitation on the proton beam core. In addition, these
measurements provide an experimental basis to guide the design and
tolerances on the residual HEL fields at the location of the beam
core, in case resonant excitation is needed for HL-LHC.

Two experiments were conducted, one in 2016~\cite{resexmd2016} and one
in 2017~\cite{resexmd2017}. Beam and machine parameters are summarized
in Table~\ref{tab:md_param}.

During the experiments, losses were measured with the fast beam
current transformers (FBCTs). Transverse beam profiles and emittances
were provided by the beam synchrotron radiation telescope (BSRT). All
instruments were capable of delivering bunch-by-bunch data. The data
analysis of the BSRT profiles is quite involved. In this paper, we
focus on the direct comparison of the resulting profiles. A detailed
description of the profile analysis can be found in
Ref.~\cite{bsrtprofinj}, with individual experiments reported in
Refs.~\cite{resexmd2016, resexmd2017}.

The choice of excitation patterns for the experiments was guided by
losses and emittance growths calculated in numerical tracking
simulations, described below. It was chosen to study experimentally
the two pulsing patterns with the largest calculated effects on the
beam ($7^{\mathrm{th}}$- and $10^{\mathrm{th}}$-turn pulsing), one
pattern with no effect ($8^{\mathrm{th}}$-turn pulsing), and the
random excitation. In order to quantify the reproducibility of the
results under different machine configurations, one pulsing pattern
($7^{\mathrm{th}}$-turn pulsing) was tried first in~2016 and then
repeated in~2017.

\subsection{Excitations with the transverse damper and filling schemes}
\label{sec:adt}

The primary function of the transverse feedback system in LHC, also
referenced to as ADT, is to provide injection oscillation damping and
actively damp the coupled bunch instabilities driven by the machine
impedance. The main building blocks of the ADT system are stipline
pickups at the position Q7 and Q9 at Point~4 of the LHC, connected to
the beam position measurement modules at the surface, the digital
signal processing modules (mDSPU) and a set of tetrode power
amplifiers feeding electrostatic kickers in the RF zone of the
LHC~\cite{adt_sum_2008,adt_sum_2011}. Though originally designed to
damp the undesired transverse activities only, the damper is being
routinely used for sophisticated beam excitations as well thanks to
the state of the art digital signal processing hardware. These include
among others abort and injection gap cleaning, single bunch excitation
for tune and linear coupling measurements and special modes of
operation for dedicated experiments at the LHC.

@@

The ADT kicker deflection angle is calculated from the kicker geometry
and the excitation voltage on the deflection plates. The excitation
voltage used in the experiments is a complicated waveform, generated
by a very complex hardware (digital signal processor, chain of
transmission lines, low and high power amplifiers). The peak
excitation voltage (and subsequently the kick) seen by the beam can be
only indirectly measured using the probes on each kicker. The measured
values are therefore believed to be known with an error margin of
10-15\% for the experiments in 2017 and 50\% in 2016. The estimate for
2017 is based on only a single measurement and in the future and with
more time available the determination of the error would definitely
have to be improved. The error in 2016 is estimated from the
transverse damper operational parameters. Due to lack of the dedicated
machine time no precision calibration was performed. The maximum kick
strength reachable with the ADT at injection without risking
saturation is 96~nrad. As an example Fig.~\ref{fig:fill_meas} shows
the oscillation amplitude of each bunch in both H and V planes
captured by the LHC real time transverse activity monitor during the
2017 experiments and an excitation with $8^{\mathrm{th}}$ turn
pulsing. The measured activity is well in line with the expected
excitation pattern.
